{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-02-08 thesis/landslide.tex
write abt time travel and friends

\subsection{Scheduler}
\subsection{Memory analysis}
\subsection{Machine state manipulation}
\subsection{State space traversal}
% time travel, explore, arbiter
\subsection{Bug-finding output}

@@

\begin{itemize}
	\item If a function is preempted at its beginning or end
		such that its corresponding base pointer is missing from the base pointer chain,
		Landslide will find its ``hidden'' frame and include it in the stack trace in the following cases.%
		\footnote{Note that in such cases,
		most other debuggers' stack tracers will be missing not the name of the interrupted function,
		but the name of the function which called that function,
		because it's the former's stack frame which should enable the debugger
		to find the pushed return address for the latter.}
		\begin{itemize}
			\item If the last pushed return address is at offset 0 into the body of its containing function,
				Landslide will find the next pushed return address at {\tt esp+0}.
			\item If as above but the function is the page fault handler, at {\tt esp+4}.
			\item If the return address is at offset 1 and the previous instruction was {\tt push ebp},
				Landslide will find the next pushed return address at {\tt esp+4}.
			\item If the return address is a multiple of 4 offset
				and all previous instructions are of the form {\tt mov m32,r32},
				Landslide will find the next pushed return address at {\tt esp+0}.
				(This is common in student hand-written assembly functions.)
		\end{itemize}
	\item If the instruction at a pushed return address is a {\tt pop} or {\tt popa},
		Landslide will search for the next non-{\tt pop}({\tt a}) instruction,
		and if it's {\tt ret} or {\tt iret},
		treat the function as a system call wrapper
		(which tend not to preserve the base pointer chain)
		and find the next return address above where all those registers were pushed.
	\item If a return address was pushed during a fault or interrupt
		(determined by checking for the {\tt iret} opcode or the page fault wrapper special case mentioned above),
		Landslide will read the iret block to determine whether a stack switched happened
		and if so what {\tt esp} used to be.
	\item Landslide runs the tortoise/rabbit algorithm to detect cyclic {\tt ebp} chains and terminate after the first time around.
	\item Two other Pebbles-specifc special cases described in \sect{\ref{sec:landslide-pebblesspecifics}}.
\end{itemize}

@@

% features:
% noreturn functions, when last instr is a call, will avoid printing the next function in the binary
% various function eip offsets
% kernel to user mode swinches

@@

The preemption traces, described and exemplified in \sect{\ref{sec:landslide-bugreport}}, are generated by {\tt found\_a\_bug.c},
in cooperation with {\tt save.c}.
Whenever {\tt save.c} creates a preemption point, it captures a stack trace of the current thread at the point it was interrupted, and saves it in the preemption point tree.
{\tt found\_a\_bug.c} then traverses the current branch of the tree,
potentially producing both console output and HTML output (controlled by the {\tt HTML\_PRINTF} macro family).
It should be invoked by the {\tt FOUND\_A\_BUG} macro defined in {\tt found\_a\_bug.h},
or by {\tt FOUND\_A\_BUG\_HTML\_INFO}, which also allows the caller to specify a callback to print extra details (such as use-after-free stack traces) in the bug report.

>>>>

\subsection{Scheduler}
\subsection{Memory analysis}
\subsection{Machine state manipulation}
\subsection{State space traversal}

Traversal of the state space is implemented in three parts:
first, identifying preemption points when first encountered and selecting which thread to run for its first execution, in {\tt arbiter.c},
second, selecting which preemption point to backtrack to after completing an execution
and which thread to ``have switched to'' %
% TODO: post-revisions; uncomment; change " %" above to "%"
%\footnote{willan on-having switched to \cite{hhgg-reu}}
instead, in {\tt explore.c},
and third, rewinding the machine state to implement said backtracking,
in {\tt timetravel.c} (Bochs version) and {\tt timetravel-simics.c} (Simics version).

\subsubsection{Arbiter}

The arbiter (named after the corresponding component of dBug \cite{dbug-ssv})
is responsible for checking which code locations during execution should be identified as preemption points
({\tt arbiter\_interested()}),
and thereupon for choosing whether to keep running the current thread or to preempt and switch to a new one
({\tt arbiter\_choose()}).
Its behaviour in the former case is configured by the options listed in \sect{\ref{sec:landslide-dynamicconfig}},
and in the latter case by the options listed in \sect{\ref{sec:landslide-staticconfig}}.
For example, {\tt EXPLORE\_BACKWARDS} is interpreted here;
if set, it will cause Landslide to always preempt and switch threads the first time it encounters each new preemption point.%
\footnote{Another secret option, {\tt CHOOSE\_RANDOMLY},
also exists here to randomize whether to ``explore backwards''
(choosing independently at each preemption point, resulting in an overall unpredictable exploration order).
% TODO: decide whether you're gonna put those estimate graphs in here
%% I
It's not exposed to {\tt config.landslide} but rather the user must edit it in {\tt arbiter.c} directly,
whereupon the probability may also be adjusted via {\tt numerator} and {\tt denominator}.}

\subsubsection{Explorer}

Landslide invokes the explorer at the end of each execution of the test case,
which analyzes the current branch of the interleaving state space tree
to figure out which alternate branch to try executing next.
Its contents are largely algorithmic rather than architectural
and hence further described in \sect{\ref{sec:landslide-dpor}} and \sect{\ref{sec:landslide-icb}}.

\subsubsection{Time Travel}

After the explorer picks a past point of program preemption, % sorry this started out like that and i just had to roll with it
Landslide collaborates with the simulator to revert the machine state to that point before switching to the desired thread.
The Simics version is merely a bunch of wrapper glue code around the {\tt set-bookmark} and {\tt skip-to} backtracking commands.
Bochs however does not support backtracking, so I instead use {\tt fork()} to get Linux to copy the machine state for me.

The big issue to note here is that, while the simulation state should be completely reverted,
parts of Landslide's state (e.g., scheduler runqueues, thread action flags)
should likewise be reverted to mirror the change in program state,
while others (tagged ancestor branches from DPOR, state space estimates) should be preserved from branch to branch.
In Simics, I simply copy every data structure of the former case ({\tt copy\_sched()} and friends in {\tt save.c}),
leaving those of the latter undisturbed across backtracks.%
\footnote{Simics actually wants to save/restore all its modules' internal state on its own,
offering an attribute set/get API for modules to expose such state (used for other purposes in {\tt simics\_glue.c}),
but doing deep copies of data structures this way would be more trouble than it's worth.}

In Bochs, {\tt fork()} automatically copies everything, so the reverse holds:
all data of the latter case must whenever updated be propagaged to all {\tt fork()}ed children processes explicitly.
I worried while implementing this that I might miss a case, or that future updates to the code could easily forget this step,
resulting undoubtedly in state corruption bugs which to diagnose would be a thesis in their own right,
so I enlisted help from my compiler via the oft-ridiculed {\tt const}.
Every preemption point node in the execution tree ({\tt tree.h}),
each of whose state is kept generally read-only,
and all modifications must go through {\tt modify\_hax()} ({\tt timetravel.h}) using a modification callback,
which internally casts away the {\tt const}, performs the requested modification,
and also messages all relevant child processes to perform the same ({\tt timetravel\_set()} in {\tt timetravel.c}).
The {\tt const} is absolutely, inviolably, not to be casted away, at the sacred cost of what little type safety C offers.%
\footnote{Of course this would be followed by a footnote describing the one place where I cast it away anyway,
{\tt mem\_check\_shared\_access()} in {\tt memory.c};
why it's ok is documented in an {\tt XXX} comment.}
Thence the typechecker enforces that all exploration-related state is properly propagated while scheduler state is automatically reverted.

% time travel, explore, arbiter
\subsection{Bug-finding output}

@@

\begin{itemize}
	\item If a function is preempted at its beginning or end
		such that its corresponding base pointer is missing from the base pointer chain,
		Landslide will find its ``hidden'' frame and include it in the stack trace in the following cases.%
		\footnote{Note that in such cases,
		most other debuggers' stack tracers will be missing not the name of the interrupted function,
		but the name of the function which called that function,
		because it's the former's stack frame which should enable the debugger
		to find the pushed return address for the latter.}
		\begin{itemize}
			\item If the last pushed return address is at offset 0 into the body of its containing function,
				Landslide will find the next pushed return address at {\tt esp+0}.
			\item If as above but the function is the page fault handler, at {\tt esp+4}.
			\item If the return address is at offset 1 and the previous instruction was {\tt push ebp},
				Landslide will find the next pushed return address at {\tt esp+4}.
			\item If the return address is a multiple of 4 offset
				and all previous instructions are of the form {\tt mov m32,r32},
				Landslide will find the next pushed return address at {\tt esp+0}.
				(This is common in student hand-written assembly functions.)
		\end{itemize}
	\item If the instruction at a pushed return address is a {\tt pop} or {\tt popa},
		Landslide will search for the next non-{\tt pop}({\tt a}) instruction,
		and if it's {\tt ret} or {\tt iret},
		treat the function as a system call wrapper
		(which tend not to preserve the base pointer chain)
		and find the next return address above where all those registers were pushed.
	\item If a return address was pushed during a fault or interrupt
		(determined by checking for the {\tt iret} opcode or the page fault wrapper special case mentioned above),
		Landslide will read the iret block to determine whether a stack switched happened
		and if so what {\tt esp} used to be.
	\item If a return address's offset into its containing function is 0,
		and the last instruction in the preceding function (binary-wise) is a {\tt call},
		Landslide will recognize it as a {\tt noreturn} tail-call, and print the correct function name.%
		\footnote{Normally Landslide reports function/line number information for the return address as-is,
		which indicates the next line of code after the relevant call rather than the call itself.}
	\item Landslide runs the tortoise/rabbit algorithm to detect cyclic {\tt ebp} chains and terminate after the first time around.
	\item Two other Pebbles-specifc special cases described in \sect{\ref{sec:landslide-pebblesspecifics}}.
\end{itemize}

@@

@@

The preemption traces, described and exemplified in \sect{\ref{sec:landslide-bugreport}}, are generated by {\tt found\_a\_\allowbreak{}bug.c},
in cooperation with {\tt save.c}.
Whenever {\tt save.c} creates a preemption point, it captures a stack trace of the current thread at the point it was interrupted, and saves it in the preemption point tree.
{\tt found\_a\_bug.c} then traverses the current branch of the tree,
potentially producing both console output and HTML output (controlled by the {\tt HTML\_PRINTF} macro family).
It should be invoked by the {\tt FOUND\_A\_BUG} macro defined in {\tt found\_a\_bug.h},
or by {\tt FOUND\_A\_BUG\_HTML\_INFO}, which also allows the caller to specify a callback to print extra details (such as use-after-free stack traces) in the bug report.
